\section{First Authored Publications}
\label{sec:first_authored_publications}

\begin{itemize}
	\item \textbf{Carse, J}. \& McKenna, S. (2019), Active learning for patch-based digital pathology using convolutional neural networks to reduce annotation costs, \textit{in} 'European Congress on Digital Pathology', Springer, pp. 20-27. 
	
	\item \textbf{Carse, J}., Carey, F. \&  McKenna, S. (2021), Unsupervised representation learning from pathology images with multi-directional contrastive predictive coding, \textit{in} '$18^{\text{th}}$ International Symposium on Biomedical Imaging', IEEE, pp. 1254-1258.
	
	\item \textbf{Carse, J}., S\"{u}veges, T., Hogg, S., Trucco, E., Proby, C., Fleming, C. \& McKenna, S. (2021), Robust selective classification of skin lesions with asymmetric costs, \textit{in} 'Uncertainty for Safe Utilization of Machine Learning in Medical Imaging, and Perinatal Imaging, Placental and Preterm Image Analysis', Springer, pp. 112-121.
	
	\item \textbf{Carse, J}., Alvarez Olmo, A. \& McKenna, S. (2022) Calibration of deep medical image classifiers: An empirical comparison using dermatology and histopathology datasets, \textit{in} 'International Workshop on Uncertainty for Safe Utilization of Machine Learning in Medical Imaging', Springer, pp. 89-99.
\end{itemize}



\section{Contributed Publications}
\label{sec:contributed_publications}

\begin{itemize}
	\item Chin, G.,  S\"{u}veges, T., \textbf{Carse, J}., Butt, S., Muthiah, S., Morton, C., Trucco, E., Proby, C., McKenna, S. \& Fleming, C. (2022), 'Prepare to succeed: Real-world image datasets for artificial intelligence in skin cancer triage', \textit{British Journal of Dermatology} \textbf{187}(Supplement 1), 125.
\end{itemize}



\section{Awards}
\label{sec:awards}

\begin{itemize}
	\item Best Paper - University of Dundee Computing PhD Symposium 2019
	\item Best PhD Blitz Presentation - SINAPSE Annual General Meeting 2019
	\item Best Reviewer - University of Dundee Computing PhD Symposium 2022
\end{itemize}